---
title: "day-08"
format: html
editor: visual
---

#Maya McCain
#February 22, 2025
#ESS 330 Daily Exercise 8

#1. Read in the COVID-19 data
```{r}
library(tidyverse)
url = 'https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties-recent.csv'
covid = read_csv(url)
library(dplyr)
```
#2. Create a new data.frame using the available state.abb, state.name, state.region objects in base R. Be intentional about creating a primary key to match to the COVID data!
```{r}
covid_new = data.frame(region = state.region,
                state_ID = state.abb,
                state = state.name)
```

#3. Join your new data.frame to the raw COVID data. Think about right, inner, left, or full join… 
```{r}
joined_covid <- full_join(covid_new, covid, by = "state")
```

#4. split-apply the joined data to determine the daily, cumulative, cases and deaths for each region
```{r}
regional_summary <- joined_covid %>%
  group_by(region, date) %>%  
  summarise(
    daily_cases = sum(cases, na.rm = TRUE),  
    daily_deaths = sum(deaths, na.rm = TRUE)
  ) %>%
  arrange(region, date) %>%  
  group_by(region) %>%
  mutate(
    cumulative_cases = cumsum(daily_cases), 
    cumulative_deaths = cumsum(daily_deaths)
  )

head(regional_summary)

```

#5. Pivot your data from wide format to long
```{r}
regional_long <- regional_summary %>%
  pivot_longer(
    cols = c(daily_cases, daily_deaths, cumulative_cases, cumulative_deaths),  
    names_to = "metric",   
    values_to = "value"    
  )

head(regional_long)

```

#6. Plot your data in a compelling way (setup, layers, labels, facets, themes)
```{r}
ggplot()
cases_deaths_plot <- ggplot(data = regional_long, 
               aes(x = date, y = value, color = region)) +
  geom_line() +
  labs(title = "Cumulative COVID-19 Cases & Deaths by USA Region",
       x = "Date",
       y = "Value",
       color = "Region") +
  facet_wrap(~metric, scales = "free_y") +
  theme_minimal()
print(cases_deaths_plot)
```

#7. Save the image to your img directory with a good file name and extension!
